# Covid 19 Data Visualization

---
title: "Covid_19"
author: "Lin Myat Ko"
date: "2023-09-07"
output: pdf_document
---

## Importing Libraries

```{r}
library(tidyverse)
library(dplyr)
library(zoo)
```

## Data Loading

We received the data from COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.

Here is the link for the data. <https://github.com/CSSEGISandData/COVID-19>

```{r}
df_confirmed = read_csv("./time_series_covid19_confirmed_global.csv")
head(df_confirmed)
```

## Data Visualization

The data contains many countries. So, to present the data effectively, I will select a few countries to visualize. I want to compare the G7 countries.

```{r}
selected_countries <- c("Canada", "France", "Germany", "Italy", "Japan", "United Kingdom", "US")

# Filtering the data by selected countries
g7_df_confirmed <- df_confirmed %>% filter(`Country/Region` %in% selected_countries)

g7_df_confirmed <- g7_df_confirmed %>% rename(country = `Country/Region`)

```

I want to draw simple line plots of the aggregate data. The data have provinces and states but I am more interested to see the country as a whole to compare with other countries. In order to do that, we first need to aggregate the cases by the country name. Before we aggregate, we will drop columns for "Province/State", "Lat", "Long" since they are not going to be useful for our first visualization.

```{r}
g7_df_confirmed[c("Province/State", "Lat", "Long")] = list(NULL)
g7_df_confirmed_aggregate <- aggregate(. ~ country, g7_df_confirmed, sum)

```

```{r}
head(g7_df_confirmed_aggregate)
```

Create pivot to be visualized as line plot

```{r}
g7_df_confirmed_aggregate_pivot <- g7_df_confirmed_aggregate %>%
  pivot_longer(
    cols = -country,
    names_to = "date",
    values_to = "cases"
  )

# Covert the day column to date object
g7_df_confirmed_aggregate_pivot$date <- as.Date(g7_df_confirmed_aggregate_pivot$date, format="%m/%d/%y")
head(g7_df_confirmed_aggregate_pivot)
```

```{r fig.width=12, fig.height=4}
df_viz <- g7_df_confirmed_aggregate_pivot %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

ggplot(df_viz, aes(x = date_interval, y = total_cases, color = country)) +
  geom_line() +
  labs(title = "Total Cases over the time",
       x = "Date",
       y = "Total Cases") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y")
```

Infectious diseases spread by exponential rate. So, let's try to visualize with log scale.

```{r fig.width=12, fig.height=4}
df_viz <- g7_df_confirmed_aggregate_pivot %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

ggplot(df_viz, aes(x = date_interval, y = total_cases, color = country)) +
  geom_line() +
  labs(title = "Total Cases over the time",
       x = "Date",
       y = "Total Cases") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") +
  scale_y_log10()
```

### Out of the G7, Japan had a significant rise in confirmed cases starting from Feb 2022.

## Moving Average Plot for Japan

I want to do moving averages to see if there is something interesting in the data.

```{r fig.width=20, fig.height=12}
df_japan <- g7_df_confirmed_aggregate_pivot %>% 
  filter(country == "Japan")

# Sort and calculate moving averages
df_moving_avg <- df_japan %>%
  arrange(date) %>%
  mutate(
    moving_avg_7day = rollmean(cases, 7, align = "right", fill = NA),
    moving_avg_30day = rollmean(cases, 30, align = "right", fill = NA),
    moving_avg_100day = rollmean(cases, 100, align = "right", fill = NA)
  )

# Create the plot
ggplot(df_moving_avg, aes(x = date, y = cases)) +
  geom_line(aes(color = "Actual"), alpha = 0.5) +
  geom_line(aes(y = moving_avg_7day, color = "7-day MA")) +
  geom_line(aes(y = moving_avg_30day, color = "30-day MA")) +
  geom_line(aes(y = moving_avg_100day, color = "100-day MA")) +
  scale_color_manual(values = c("Actual" = "black", "7-day MA" = "red", "30-day MA" = "green", "100-day MA" = "blue")) +
  labs(title = "Cases and Moving Averages for ABC",
       x = "Date",
       y = "Cases",
       color = "Legend") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y")
```

## Important Deaths global dataset

```{r}
df_death = read_csv("./time_series_covid19_deaths_global.csv")
head(df_death)

```

Again, we will looking only at the G7 countries.

```{r}
selected_countries <- c("Canada", "France", "Germany", "Italy", "Japan", "United Kingdom", "US")

# Filtering the data by selected countries
g7_df_death <- df_death %>% filter(`Country/Region` %in% selected_countries)

g7_df_death <- g7_df_death %>% rename(country = `Country/Region`)
head(g7_df_death)
View(g7_df_death)
```

```{r}
g7_df_death[c("Province/State", "Lat", "Long")] = list(NULL)
g7_df_death_aggregate <- aggregate(. ~ country, g7_df_death, sum)
head(g7_df_death_aggregate)

```

```{r}
g7_df_death_aggregate_pivot <- g7_df_death_aggregate %>%
  pivot_longer(
    cols = -country,
    names_to = "date",
    values_to = "deaths"
  )

# Covert the day column to date object
g7_df_death_aggregate_pivot$date <- as.Date(g7_df_death_aggregate_pivot$date, format="%m/%d/%y")
head(g7_df_death_aggregate_pivot)
```

```{r fig.width=12, fig.height=4}
df_viz <- g7_df_death_aggregate_pivot %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_deaths = sum(deaths, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

ggplot(df_viz, aes(x = date_interval, y = total_deaths, color = country)) +
  geom_line() +
  labs(title = "Total Death over the time",
       x = "Date",
       y = "Total Deaths") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y")
```

#### US data is much more than other countries. From this visualization, we might assume that other countries can keep fatality under control. So, we will plot with US to see the real trends.

```{r fig.width=12, fig.height=4}
df_without_us <- g7_df_death_aggregate_pivot %>% 
  filter(country != "US")

df_viz <- df_without_us %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_deaths = sum(deaths, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

ggplot(df_viz, aes(x = date_interval, y = total_deaths, color = country)) +
  geom_line() +
  labs(title = "Total Death over the time",
       x = "Date",
       y = "Total Deaths") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y")
```

#### Visualization With US

### Correlation between confirmed and deaths for US

```{r fig.width=12, fig.height=4}
df_us_confirmed <- g7_df_confirmed_aggregate_pivot %>% 
  filter(country == "US")

df_viz <- df_us_confirmed %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

plot <- ggplot(df_viz, aes(x = date_interval, y = total_cases, color = "Cases")) +
  geom_line() +
  labs(title = "Total Cases and Deaths over the time",
       x = "Date",
       y = "Total Cases") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") +
  scale_y_log10()

df_us_death <- g7_df_death_aggregate_pivot %>% 
  filter(country == "US")

df_viz2 <- df_us_death %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_deaths = sum(deaths, na.rm = TRUE))

df_viz2$date_interval <- as.Date(as.character(df_viz2$date_interval))

plot + 
  geom_line(data = df_viz2, aes(x = date_interval, y = total_deaths, color = "Deaths"))

```

### Correlation between confirmed and deaths for Japan

```{r fig.width=12, fig.height=4}
df_us_confirmed <- g7_df_confirmed_aggregate_pivot %>% 
  filter(country == "Japan")

df_viz <- df_us_confirmed %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE))

df_viz$date_interval <- as.Date(as.character(df_viz$date_interval))

plot <- ggplot(df_viz, aes(x = date_interval, y = total_cases, color = "Cases")) +
  geom_line() +
  labs(title = "Total Cases over the time",
       x = "Date",
       y = "Total Cases") +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") +
  scale_y_log10()

df_us_death <- g7_df_death_aggregate_pivot %>% 
  filter(country == "Japan")

df_viz2 <- df_us_death %>%
  group_by(country, date_interval = cut.Date(date, breaks = "10 day")) %>%
  summarise(total_deaths = sum(deaths, na.rm = TRUE))

df_viz2$date_interval <- as.Date(as.character(df_viz2$date_interval))

plot + 
  geom_line(data = df_viz2, aes(x = date_interval, y = total_deaths, color = "Deaths"))
```
