# NYPD Shooting Incident Data

---
title: "NYPD Shooting Incident Data"
author: "Lin Myat Ko"
date: "2023-09-06"
output: pdf_document
---

```{r}
library(tidyverse)
```

## Data Loading

```{r}
df <- read_csv("./NYPD/NYPD_Shooting_Incident_Data__Historic_.csv")
```

## Summary Statistics

```{r}
summary(df)
```

## Data Visualization

### 1. Most Risky Hour of the Day

Let's explore the OCCUR TIME. I want to find out which are the riskiest hours. The intuition is it is safer in the daytime at night.

```{r fig.width=12, fig.height=8}
breaks_hours <- seq(0, 23, 1)
labels_hours <- c("12 AM", paste0(seq(1, 11, 1), " AM"), "12 PM", paste0(seq(1, 11, 1), " PM"))


df$hour <- hour(hms(df$OCCUR_TIME))

ggplot(df, aes(x = hour)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.7, color = "black") +
  labs(title = "Most risky time of the day",
       x = "Time of the day",
       y = "Shootings") +
  scale_x_continuous(breaks = breaks_hours, labels = labels_hours) +
  theme_minimal()

```

So, our intuition is backed by the data.

### 2. Day of the Week

Next, I want to determine the safest day of the week.

This is going to be tricky because we do not have the day of the week in data. However, we can calculate day of the week from date.

```{r fig.width=12, fig.height=8}
df$weekday <- wday(as.Date(df$OCCUR_DATE, format = "%m/%d/%Y"), label = TRUE)

df$weekday <- factor(df$weekday, levels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))

ggplot(df, aes(x = weekday)) +
  geom_bar(fill = "blue", alpha = 0.7, color = "black") +
  labs(title = "Shootings by day of the week",
       x = "Day of the week",
       y = "Frequency") +
  theme_minimal()


```

This is interesting to see Wednesday is the safest compared to the weekend.

### 3. Distribution of races

Let us explore the races involved. I will print the statistics.

```{r fig.width=12, fig.height=8}
ggplot(df, aes(x = VIC_RACE)) +
  geom_bar(fill = "red", alpha = 0.7, color = "black") +
  labs(title = "Victim by race",
       x = "Race",
       y = "Number") +
  theme_minimal()
```

#### The numbers are too different so I will use LOG SCALE to compare victim and **perpetrator** distributions.

```{r fig.width=12, fig.height=8}
ggplot(df, aes(x = VIC_RACE)) +
  geom_bar(fill = "red", alpha = 0.7, color = "black") +
  labs(title = "Victim by race (LOG SCALE)",
       x = "Race",
       y = "LOG SCALE") +
  scale_y_log10() +
  theme_minimal()
```

```{r fig.width=12, fig.height=8}
ggplot(df, aes(x = PERP_RACE)) +
  geom_bar(fill = "black", alpha = 0.7, color = "black") +
  labs(title = "Perpetrator by race (LOG SCALE)",
       x = "Race",
       y = "LOG SCALE") +
  scale_y_log10() +
  theme_minimal()
```

I cannot help but feel that data is biased.

### 4. Distributions of Victim Races by District

Let's explore the distributions of races by districts are equal.

```{r fig.width=12, fig.height=8}
library(ggplot2)

ggplot(df, aes(x = VIC_RACE)) +
  geom_histogram(stat = "count", aes(fill = VIC_RACE), position = "dodge") +
  facet_wrap(~ BORO) +
  scale_y_log10() +
  ylab("Frequency (Log Scale)") +
  ggtitle("Distribution of Races by District (Log Scale)") +
  theme_minimal() +
  theme(axis.text.x=element_blank())

```

We can see that all districts are more of less similarly distributed.

### 5. Distributions of Victim Age Group by District

```{r}
df <- df[!(df$VIC_AGE_GROUP == "UNKNOWN" | df$VIC_AGE_GROUP == 1022),]
ggplot(df, aes(x = VIC_AGE_GROUP)) +
  geom_histogram(stat = "count", aes(fill = VIC_AGE_GROUP), position = "dodge") +
  facet_wrap(~ BORO) +
  scale_y_log10() +
  ylab("Frequency (Log Scale)") +
  ggtitle("Distribution of Age Group by District (Log Scale)") +
  theme_minimal() +
  theme(axis.text.x=element_blank()) 
```

They all seem to distribute similarly.

## Modelling

I want to do logistic regression statistical murder flag. I will handpick the features I want to explore.

OCCUR_TIME, PRECINT, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, VIC_AGE_GROUP, VIC_SEX, VIC_RACE

### Data Cleaning

```{r}
model_df <- df[, c("OCCUR_TIME", "PRECINCT", "PERP_AGE_GROUP", "PERP_SEX", "PERP_RACE", "VIC_AGE_GROUP", "VIC_SEX", "VIC_RACE", "STATISTICAL_MURDER_FLAG")]

model_df <- na.omit(model_df)
model_df <- model_df[!apply(model_df == "UNKNOWN", 1, any), ]

model_df$TIME_OF_DAY <- factor(hour(hms(model_df$OCCUR_TIME)))
```

```{r}
model_df[] <- lapply(model_df, factor)
```

```{r}
table(model_df$STATISTICAL_MURDER_FLAG)
```

But the data is imbalanced so, I will sample 3000 from FALSE.

I will handpick the features I want to explore.

### Feature Selection

OCCUR_TIME, PRECINT, PERP_AGE_GROUP, PERP_SEX, PERP_RACE, VIC_AGE_GROUP, VIC_SEX, VIC_RACE

```{r}
sampled_df <- model_df %>%
  group_by(STATISTICAL_MURDER_FLAG) %>%
  sample_n(3000, replace = FALSE) %>%
  ungroup()

```

```{r}
set.seed(123)
sample <- sample(c(TRUE, FALSE), nrow(sampled_df), replace=TRUE, prob=c(0.8,0.2))
train  <- sampled_df[sample, ]
test   <- sampled_df[!sample, ]


```

```{r}
model = glm(STATISTICAL_MURDER_FLAG ~ PRECINCT + PERP_AGE_GROUP + PERP_SEX + PERP_RACE + VIC_AGE_GROUP + VIC_SEX + VIC_RACE + TIME_OF_DAY, data = train, family = binomial)
```

```{r}
predictions = predict(model, newdata = test, type = "response")
```

```{r}
predicted_labels = ifelse(predictions > 0.5, TRUE, FALSE)

library(caret)
confusionMatrix(as.factor(predicted_labels), as.factor(test$STATISTICAL_MURDER_FLAG))
```

### Conclusion

So, the model accuracy is bad. So, I think it's safe to conclude that the NYPD determine statistical murder flag by some kinds of private data. In the future, I will research more about how data scientists are looking at this kind of data.
